### 2018-01-05 2018/images/pic2.tex
todo corr done

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}

%Angle Definitions
%-----------------

%start tikz picture, and use the tdplot_main_coords style to implement the display
%coordinate transformation provided by 3dplot
\tdplotsetmaincoords{40}{130}

\begin{tikzpicture}[scale=5,tdplot_main_coords]

\coordinate (O) at (0,0,0);
\coordinate (xc) at (1,0);
\coordinate (yc) at (0,2);
\coordinate (un) at (0,0,4);

\tdplotsetcoord{S}{2}{30}{0}  % икс
\tdplotsetcoord{P}{4}{30}{90} % игрек
\tdplotsetcoord{Z}{3.4641}{0}{0} % 3.4641 = \sqrt{4^2 - 2^2}

%\draw[color=white, fill=gray!10] (xc) to[out=90,in=90] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
%\draw[color=white, fill=blue!10] (yc) to[out=90,in=90] (P) to[out=90,in=0] (un) -- (O) -- cycle;

\draw[color=white, fill=gray!5] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
\draw[color=white, fill=blue!5] (P) to[out=90,in=0] (un) -- (O) -- cycle;

%\draw[color=white, fill=gray!10, pattern=dots] (S)  to[out=90,in=180]  (un) -- (O) -- cycle;
%\draw[color=white, fill=blue!10, pattern=north east lines] (P) to[out=90,in=0] (un) -- (O) -- cycle;

\draw[thick,->] (O) -- (xc) node[anchor=north east]{$x - \bar{x} \cdot \mathbf{1}$};
\draw[thick,->] (O) -- (yc) node[anchor=north west]{$y - \bar{y} \cdot \mathbf{1}$};
\draw[thick,->] (O) -- (un) node[anchor=south]{$\mathbf{1}$};

\tdplotdrawarc{(O)}{0.1}{0}{90}{anchor=north}{$\phi$}

\draw[-stealth,thick] (O) -- (S) node[anchor=north east]{$x$};
\draw[-stealth,thick] (O) -- (P) node[anchor=north west]{$y$};

\draw[dashed] (S) -- (Sx);
\draw[dashed] (P) -- (Py);

\tkzMarkRightAngle[size=0.05](P,yc,O);
\tkzMarkRightAngle[size=0.05](S,xc,O);

\end{tikzpicture}

\end{document}