### 2018-01-07 2018/images/corr_const_br_up.tex
corr and r^2

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}

%Angle Definitions
%-----------------

%start tikz picture, and use the tdplot_main_coords style to implement the display
%coordinate transformation provided by 3dplot
\tdplotsetmaincoords{40}{120}

\begin{tikzpicture}[tdplot_main_coords]

\coordinate (O) at (0,0,0);
\coordinate (xc) at (2.1,-2.1);
\coordinate (xcp) at (0.84,-0.84);
\coordinate (yc) at (-0.9,3);
\coordinate (un) at (0,0,6);
\coordinate (unt) at (0,0,2.9);

\tdplotsetcoord{S}{10}{35}{0}  % икс
\tdplotsetcoord{R}{4}{35}{0} % разложение
\tdplotsetcoord{P}{7}{30}{90} % игрек
\tdplotsetcoord{T}{6.5}{20}{0} % икс с тильдой

\draw[->] (O) -- (un) node[anchor=east]{$\mathbf{1}$};
\draw[->] (O) -- (S) node[anchor=south east]{$x$};
\draw[->, gray] (O) -- (P) node[anchor=north west, gray]{$y$};
\draw[->] (O) -- (T) node[anchor=south east]{$\tilde x$};

\draw (0,0.5) ellipse (4cm and 1cm) node[anchor=west, right=2.7cm]{\tiny$Lin^{\perp}(\mathbf{1})$};

\draw[->, gray] (O) -- (yc) node[anchor=north east]{\small $y^c$};
\draw[->] (O) -- (xc) node[anchor=north east]{\small $x^c$};
\draw[dashed, gray] (P) -- (yc);
\draw[dashed] (S) -- (xc);

\draw[->, dashed, blue] (unt) -- (T);
\draw[->, dashed, red] (R) -- (T);
\draw[->, blue] (O) -- (R) node[anchor=north east]{\footnotesize$\beta x$};
\draw[->, red] (O) -- (unt) node[anchor=west]{\footnotesize$\alpha \mathbf{1}$};

\draw[dashed, blue] (R) -- (xcp);
\draw[->,dashed,blue] (O) -- (xcp) node[anchor=north]{\small$\tilde x^c$};
\tkzMarkRightAngle[size=0.15,color=blue](R,xcp,O);
\tkzDrawPoint[color=red,fill=red,size=1.8](O);

\tkzMarkRightAngle[size=0.15,color=gray](P,yc,O);
\tkzMarkRightAngle[size=0.15](S,xc,O);
\tdplotdrawarc{(O)}{0.15}{-45}{110}{anchor=north}{\tiny$\varphi$}

\end{tikzpicture}
\end{document}