### 2018-01-07 2018/images/corr_def.tex

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usetikzlibrary{calc}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}
\begin{tikzpicture}

\tkzDefPoint(0,0){O}
\tkzDefPoint(4,0){X}
\tkzDefPoint(2,3){Y}

\draw[->] (O) -- (X) node[anchor=north east]{$x$};
\draw[->] (O) -- (Y) node[anchor=east]{$y$};

\tkzMarkAngle[size=0.4](X,O,Y)
\tkzLabelAngle[pos=0.6](X,O,Y){$\varphi$}

\end{tikzpicture}
\end{document}