### 2018-01-31 2018/images/f_dist_def.tex
geometry of r.v.s

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}

%Angle Definitions
%-----------------

%start tikz picture, and use the tdplot_main_coords style to implement the display
%coordinate transformation provided by 3dplot
\tdplotsetmaincoords{40}{130}

\begin{tikzpicture}[tdplot_main_coords]

\coordinate (O) at (0,0,0);
\coordinate (un) at (0,0,4);
\coordinate (nun) at (0,0,-2.2);
\coordinate (hun) at (0,0,3.2);
\coordinate (z) at (0,3,5);
\coordinate (hz) at (-1.05,2.05,0);

\draw[->] (O) -- (un) node[anchor=east]{$L_1$};
\draw[dashed] (O) -- (nun);
\draw (0,0) ellipse (4.5cm and 1.5cm) node[anchor=east, left=3.2cm]{$L_2$};

\draw[->] (O) -- (z) node[anchor=south]{$z$};
\draw[->] (O) -- (hz) node[anchor=north]{$H_2 z$};
\draw[->] (O) -- (hun) node[anchor=east]{$H_1 z$};
\draw[dashed] (hz) -- (z);
\draw[dashed] (z) -- (hun);
\tkzMarkRightAngle[size=0.15](z,hz,O);
\tkzMarkRightAngle[size=0.15](O,hun,z);

\end{tikzpicture}
\end{document}