### 2018-02-02 2018/figures/pic3.tex

\documentclass{article}
%%%<
\usepackage{verbatim}
%%%>

\usepackage{tikz}   %TikZ is required for this to work.  Make sure this exists before the next line

\usepackage{tikz-3dplot} %requires 3dplot.sty to be in same directory, or in your LaTeX installation
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage[active,tightpage]{preview}  %generates a tightly fitting border around the work
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{2mm}

\begin{document}
\tdplotsetmaincoords{100}{110}
\begin{tikzpicture}[scale=5,tdplot_main_coords]

\coordinate (O) at (0,0,0);
\coordinate (un) at (0.75,0.75,0);
\coordinate (bx) at (-1.75,0.5,0);
\coordinate (unl) at (1.5,1.5,0);

\tdplotsetcoord{Y}{1.75}{30}{90}

\draw[->] (O) -- (bx) node[anchor=south west]{$x$};
\draw[->] (O) -- (Y) node[anchor=south west]{$y$};
\draw[->] (O) -- (un) node[anchor=south east]{$\bar{y}\cdot\mathbf{1}$};
\draw[dashed,->] (O) -- (unl) node[anchor=south west]{$\mathbf{1}$};

\draw[->] (O) -- (Yxy) node[anchor=south west]{$\hat{y}$};
\draw[dashed, color=blue] (Y) -- (Yxy) node[midway,color=black,anchor=west]{\tiny$\sqrt{RSS}$};
\draw[dashed, color=blue] (Y) -- (un) node[midway,color=black,anchor=east]{\tiny$\sqrt{TSS}$};
\draw[dashed, color=blue] (un) -- (Yxy) node[midway,color=black,anchor=north east]{\tiny$\sqrt{ESS}$};

\tkzMarkRightAngle[size=0.03](Y,Yxy,un);

\tdplotsetthetaplanecoords{110}
\tdplotdrawarc[tdplot_rotated_coords]{(un)}{0.03}{12}{110}{anchor=south west}{\footnotesize$\varphi$}

\end{tikzpicture}

\end{document}